Hydrodynamics of a Semipermeable Inextensible Membrane Under Flow and Confinement

### 2021-02-17 vesicle_code/docs/permeableVes/semipermeable_vesicle.tex
yny

The cellular membrane is a lipid bilayer membrane that is slightly permeable to water and lipid-soluble
molecules~\cite{Dick1964_JTB, FettiplaceHaydon1980_PhysRev,
DeamerBramhall1986_ChemPhysLipids, Grafmueller2019_ABLS}. 
%
%In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows against the osmotic gradient  to neutralize the osmolarity. 
%
In the red blood cell membrane, aquaporin-1 channels give rise to a much
higher water permeability coefficient of approximately $1.8\times 10^{-2}$ cm/s in
erythrocytes~\cite{YangMaVerkman2001_JBC} while the intrinsic water
permeability of a lipid bilayer membrane is in the range of
$10^{-4}$--$10^{-3}$ cm/s~\cite{ThompsonHuang1966_ANYAS,
FettiplaceHaydon1980_PhysRev, Grafmueller2019_ABLS, Dimova2020_GVB,
BhatiaRobinsonDimova2020_SoftMatt}. 
Without aquaporins or other membrane channels to promote water
permeability, the intrinsic membrane permeability is known to depend on the
temperature, lipid composition~\cite{OlbrichRawiczNeedhamEtAl2000_BJ},
and the hydration of the membrane~\cite{MarrinkBerendsen1994_JPhysChem}.
%
%The swelling/shrinking of a cell is inevitably coupled to changes in
%membrane tension and deformation. Under a strong confinement (such as
%going through a microcapillary or submicron slit),  the transmembrane
%water flow may have comparable contributions from both osmotic and
%mechanical stresses
%\cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}.  
Molecular dynamic (MD) simulations show that a stretched red blood cell
membrane porates when the membrane tension reaches the order of $2$--$4$
mN/m (see~\citet{RazizadehNikfarPaulLiu2020_BJ} and references therein),
giving rise to enhanced membrane permeability to both water and
macromolecules such as ATP. Consequently, as long as the lipid bilayer membrane
remains intact (no poration or rupture), the total amount of water inside the membrane 
is assumed constant throughout the experiment of duration
no longer than tens of minutes.
%
%membrane poration is inevitable when membrane tension is of the order
%of $2-4$ mN/m (\cite{RazizadehNikfarPaulLiu2020_BJ} and references
%therein). 
%%The CGMD results further show that a stretched RBC membrane may have pores that might contribute to transmembrane transport of both water and ATP molecules \cite{RazizadehNikfarPaulLiu2020_BJ}. 
%Thus an important outstanding question in stress-induced ATP release of
%highly deformed RBCs is if there is any transmembrane flow when RBCs
%release ATP as they experience large mechanical load due to confinement
%or large shear stress?  To address such question, the first step we
%take is to focus on the permeable water flux induced by mechanical
%stress.
%%
%%RBCs are known to release ATP under large deformation and microfluidic experiments have been used to study the correlation between RBC deformation and ATP release \cite{Wan2008_PNAS,ForsythWan2011_PNAS,ZhangShenHoganBarakatMisbah2018_BJ}.
%%
%%Why is it interesting to study the dynamics of a semipermeable membrane induced by mechanical stress/confinement?
%%
%%The osmolarity-induced water flux has been well studied, while mechanical pressure may also contribute to the permeable water flow \cite{yao-mor2017}.
%%The response of a cell to an osmotic gradient is inevitably both mechanical (tension redistribution and shape change) and chemical (water flux across the membrane).

@@

The value of $\beta$ can be estimated by first
converting the apparent permeability of a 
membrane~\cite{OlbrichRawiczNeedhamEtAl2000_BJ} to the hydraulic
conductivity $k_w$ using the conversion
in~\citet{FettiplaceHaydon1980_PhysRev}, and then scaling $k_w$ to $R_0/\mu$. We found that for most lipid bilayer membranes, $10^{-14}\le
k_w\le 10^{-12}$ m$^2$s/kg, thus the range of $\beta$ corresponding to the native water permeability of a lipid bilayer membrane (vesicle) is $10^{-11}\le
\beta\le 10^{-9}$, consistent with the value in \citet{vogl2014effect}. In the following we use a wider range of $\beta$ to elucidate the effects of permeating water flow across a
semipermeable inextensible membrane.
%

For example, we use $10^{-5}\le\beta\le10^{-2}$ in our simulations of a
semipermeable inextensible membrane in free space under various flowing
conditions. We found that the equilibrium shape of a semipermeable
membrane in free space to be independent of $\beta$. Within this range
of $\beta$ we find that $\beta$ only affects the time it takes for the
membrane to reach the equilibrium shape, which is determined by the
background flow (such as a quiescent flow, a linear shear flow and its
shear rate, and a Poiseuille flow and its maximum velocity). From these
results and our analysis of area dynamics, we suspect similar
hydrodynamics of a semipermeable vesicle (with $10^{-11}\le\beta\le
10^{-9}$). 
%
When we use $\beta\ge 0.1$, we find that the semipermeable membrane swells to maximize its water content, independent of the initial reduced 
area and the background flow.
%
%In the simulations we focus on intermediate value of $10^{-5}\le \beta\le 10^{-3}$ to investigate the hydrodynamics of a semipermeable inextensible membrane. 
%

>>>>

The cellular membrane is a lipid bilayer membrane that is slightly permeable to water and lipid-soluble
molecules~\cite{Dick1964_JTB, FettiplaceHaydon1980_PhysRev,
DeamerBramhall1986_ChemPhysLipids, Grafmueller2019_ABLS}. 
%
%In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows against the osmotic gradient  to neutralize the osmolarity. 
%
In the red blood cell membrane, aquaporin-1 channels give rise to a much
higher water permeability coefficient of approximately $1.8\times 10^{-2}$ cm/s in
erythrocytes~\cite{YangMaVerkman2001_JBC} while the intrinsic water
permeability of a lipid bilayer membrane is in the range of
$10^{-4}$--$10^{-3}$ cm/s~\cite{ThompsonHuang1966_ANYAS,
FettiplaceHaydon1980_PhysRev, Grafmueller2019_ABLS, Dimova2020_GVB,
BhatiaRobinsonDimova2020_SoftMatt}. 
Without aquaporins or other membrane channels to promote water
permeability, the intrinsic membrane permeability is known to depend on the
temperature, lipid composition~\cite{OlbrichRawiczNeedhamEtAl2000_BJ},
and the hydration of the membrane~\cite{MarrinkBerendsen1994_JPhysChem}.
%
%The swelling/shrinking of a cell is inevitably coupled to changes in
%membrane tension and deformation. Under a strong confinement (such as
%going through a microcapillary or submicron slit),  the transmembrane
%water flow may have comparable contributions from both osmotic and
%mechanical stresses
%\cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}.  
Molecular dynamic (MD) simulations show that a stretched red blood cell
membrane porates when the membrane tension reaches the order of $2$--$4$
mN/m (see~\citet{RazizadehNikfarPaulLiu2020_BJ} and references therein),
giving rise to enhanced membrane permeability to both water and
macromolecules such as ATP. Consequently, as long as the lipid bilayer membrane
remains intact (no poration or rupture), the total amount of water inside the membrane 
is assumed constant throughout the experiment of duration
no longer than tens of minutes.
%
%membrane poration is inevitable when membrane tension is of the order
%of $2-4$ mN/m (\cite{RazizadehNikfarPaulLiu2020_BJ} and references
%therein). 
%%The CGMD results further show that a stretched RBC membrane may have pores that might contribute to transmembrane transport of both water and ATP molecules \cite{RazizadehNikfarPaulLiu2020_BJ}. 
%Thus an important outstanding question in stress-induced ATP release of
%highly deformed RBCs is if there is any transmembrane flow when RBCs
%release ATP as they experience large mechanical load due to confinement
%or large shear stress?  To address such question, the first step we
%take is to focus on the permeable water flux induced by mechanical
%stress.
%%
%%RBCs are known to release ATP under large deformation and microfluidic experiments have been used to study the correlation between RBC deformation and ATP release \cite{Wan2008_PNAS,ForsythWan2011_PNAS,ZhangShenHoganBarakatMisbah2018_BJ}.
%%
%%Why is it interesting to study the dynamics of a semipermeable membrane induced by mechanical stress/confinement?
%%
%%The osmolarity-induced water flux has been well studied, while mechanical pressure may also contribute to the permeable water flow \cite{yao-mor2017}.
%%The response of a cell to an osmotic gradient is inevitably both mechanical (tension redistribution and shape change) and chemical (water flux across the membrane).  
%
%The cellular nuclear envelop (NE) consists of two lipid bilayer membranes connected via nuclear pore complexes (NPCs). As a result, NE is has a higher hydraulic conductivity
%than a lipid bilayer membrane. Even though the NPCs are sufficiently large for transport of both macromolecules and water molecules, the colloid-osmotic pressure inside
%the nucleus is highly effective in regulating the water content (see \citet{Shahin2019_AdvSci} and references therein).

@@

The value of $\beta$ can be estimated by first
converting the apparent permeability of a 
membrane~\cite{OlbrichRawiczNeedhamEtAl2000_BJ} to the hydraulic
conductivity $k_w$ using the conversion
in~\citet{FettiplaceHaydon1980_PhysRev}, and then scaling $k_w$ to $R_0/\mu$. We found that for most lipid bilayer membranes, $10^{-14}\le
k_w\le 10^{-12}$ m$^2$s/kg, thus the range of $\beta$ corresponding to the native water permeability of a lipid bilayer membrane (vesicle) is $10^{-11}\le
\beta\le 10^{-9}$, consistent with the value in \citet{vogl2014effect}. 
The cellular nuclear envelop is often treated as a filtration membrane \cite{Shahin2019_AdvSci}, and 
%Treating the cellular nuclear envelop as an ultrafiltration membrane, 
its hydraulic conductivity is reported to be order of magnitude larger \cite{Wijmans1984_JMS}:
$10^{-11}\le k_w \le 10^{-9}$m$^2$s/kg,  and the corresponding range of $\beta$ (using the same $R_0/\mu$ for scaling) is $10^{-8}\le
\beta\le 10^{-6}$.

For the simulations in \S~\ref{sec:free_space} and \S~\ref{sec:confinement} we use a wider range of $\beta$ to elucidate the effects of permeating water flow across a
semipermeable inextensible membrane.
%
For example, we use $10^{-5}\le\beta\le10^{-2}$ in our simulations of a
semipermeable inextensible membrane in free space under various flowing
conditions. We found that the equilibrium shape of a semipermeable
membrane in free space to be independent of $\beta$. Within this range
of $\beta$ we find that $\beta$ only affects the time it takes for the
membrane to reach the equilibrium shape, which is determined by the
background flow (such as a quiescent flow, a linear shear flow and its
shear rate, and a Poiseuille flow and its maximum velocity). From these
results and our analysis of area dynamics, we suspect similar
hydrodynamics of a semipermeable vesicle (with $10^{-11}\le\beta\le
10^{-9}$). 
%
When we use $\beta\ge 0.1$, we find that the semipermeable membrane swells to maximize its water content, independent of the initial reduced 
area and the background flow.
%
%In the simulations we focus on intermediate value of $10^{-5}\le \beta\le 10^{-3}$ to investigate the hydrodynamics of a semipermeable inextensible membrane. 
%
